---
output: html_document
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.width = 8, fig.height = 6)

library(dplyr)
library(ggplot2)
library(highcharter)
library(maps)
library(scales)
library(DT)

indiancities <- read.csv("../input/cities_r2.csv")

```

<center> <h1>Top 500 Indian Cities</h1> </center>

```{r}

by_state <- indiancities %>% group_by(state_name) %>% summarise(Total = n(), Population = sum(population_total), Male_Population = sum(population_male), Female_Population = sum(population_female), Male_Percent = Male_Population/Population * 100, Female_Percent = Female_Population/Population * 100, Graduates = sum(total_graduates), Male_Grads = sum(male_graduates), Female_Grads = sum(female_graduates), Grads_percent = Graduates/Population * 100, Male_Grads_Percent = Male_Grads/Male_Population * 100, Female_Grads_Percent = Female_Grads/Female_Population * 100) %>% arrange(desc(Total))


hchart(by_state, x = state_name, y = Total, type = "column", color = state_name) %>%
  hc_title(text = "States by number of cities in Top 500") %>%
  hc_add_theme(hc_theme_google())


hchart(by_state, x = state_name, value = Total, color = Total, type = "treemap")  %>% hc_add_theme(hc_theme_google()) %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "States by Number of Cities in Top 500")

```

##### Bigger states of the country have high number of large cities.


```{r}

hchart(by_state[order(-by_state$Population),], x = state_name, y = Population, type = "column", color = state_name) %>%
  hc_title(text = "States by Total Population in Top 500") %>%
  hc_add_theme(hc_theme_google())

hchart(by_state, x = state_name, value = Population, color = Population, type = "treemap")  %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "States by Total Population in Top 500 Cities") %>%
  hc_add_theme(hc_theme_google())
 
top10_states <- indiancities[indiancities$state_name %in% by_state[1:10,]$state_name,]

highchart() %>% 
  hc_add_series_boxplot(top10_states$population_total, by = top10_states$state_name,
                        name = "Total Population", color = "blue") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Top 10 States Population Distribution") %>%
  hc_add_theme(hc_theme_google())

```


##### States which are economic power-houses of the country have more population in the cities.


```{r}

hchart(by_state, x = state_name, value = Male_Percent, color = Male_Percent, type = "treemap")  %>% 
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "States by Male Population Percentage") %>%
  hc_add_theme(hc_theme_google())

```

##### Northern States of the country have high male population rate


```{r}

hchart(by_state, x = state_name, value = Female_Percent, color = Female_Percent, type = "treemap") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "States by Female Population Percentage") %>%
  hc_add_theme(hc_theme_google())

```

##### Southern and North Eastern states have high female population rate



```{r}

hchart(by_state, x = state_name, value = Grads_percent , color = Grads_percent, type = "treemap") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "States by Total Graduates Percentage in Entire Population") %>%
  hc_add_theme(hc_theme_google())

```

##### North Eastern states along with few northern states have high number of graduates in the country

```{r}

hchart(by_state, x = state_name, value = Male_Grads_Percent , color = Male_Grads_Percent, type = "treemap") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "States by Male Graduates Percentage in Male Population") %>%
  hc_add_theme(hc_theme_google())

```


##### Just like the male population rate, male graduates rate is also high in northern states of the country


```{r}

hchart(by_state, x = state_name, value = Female_Grads_Percent , color = Female_Grads_Percent, type = "treemap") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "States by Female Graduates Percentage in Female Population") %>%
  hc_add_theme(hc_theme_google())

```


##### Female graduates rate is also high in northern and few northeastern states


### Which are the Top 50 cities of the country ?

```{r}

top50_cities <- indiancities[order(-indiancities$population_total),]

top50_cities <- top50_cities[1:50,]

by_top50_cities <- indiancities[indiancities$name_of_city %in% top50_cities$name_of_city, ] %>% group_by(name_of_city) %>% summarise(Total = n(), Population = sum(population_total), Male_Population = sum(population_male), Female_Population = sum(population_female), Male_Percent = Male_Population/Population * 100, Female_Percent = Female_Population/Population * 100, Graduates = sum(total_graduates), Male_Grads = sum(male_graduates), Female_Grads = sum(female_graduates), Grads_percent = Graduates/Population * 100, Male_Grads_Percent = Male_Grads/Male_Population * 100, Female_Grads_Percent = Female_Grads/Female_Population * 100) %>% arrange(desc(Total))

```

```{r}

head(by_top50_cities)

hchart(by_top50_cities, x = name_of_city, value = Population, color = Population, type = "treemap") %>%
      hc_add_theme(hc_theme_google()) %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Top 50 Cities by Population")


```

##### Not surprisingly the capital cities of the big economic states have the biggest population.

```{r}

hchart(by_top50_cities, x = name_of_city, value = Male_Percent, color = Male_Percent, type = "treemap")  %>% 
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Male Population Percentage in Top 50 Cities") %>%
  hc_add_theme(hc_theme_google())

```



```{r}

hchart(by_top50_cities, x = name_of_city, value = Female_Percent, color = Female_Percent, type = "treemap") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Female Population Percentage in Top 50 Cities") %>%
  hc_add_theme(hc_theme_google())

```

##### Surat has the highest male population percentage among Top 50
##### Coimbatore which is second tier city of Tamil Nadu has the lowest percentage of male population among Top 50
##### When it comes to female population percentage, the plot becomes right opposite.



```{r}

hchart(by_top50_cities, x = name_of_city, value = Grads_percent , color = Grads_percent, type = "treemap") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Total Graduates Percentage in Entire Population of Top 50 Cities") %>%
  hc_add_theme(hc_theme_google())

```



```{r}

hchart(by_top50_cities, x = name_of_city, value = Male_Grads_Percent , color = Male_Grads_Percent, type = "treemap") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Male Graduates Percentage in Male Population of Top 50 Cities") %>%
  hc_add_theme(hc_theme_google())

```



```{r}

hchart(by_top50_cities, x = name_of_city, value = Female_Grads_Percent , color = Female_Grads_Percent, type = "treemap") %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Female Graduates Percentage in Female Population of Top 50 Cities") %>%
  hc_add_theme(hc_theme_google())

```



##### Graduates percentage is highest among the capital cities of the country 
##### Even by male and female graduates percentage it is highest among the capital cities.




```{r}

hchart(indiancities, "scatter", x = effective_literacy_rate_male, y = effective_literacy_rate_female, group = state_name) %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Male Literacy rate versus Female Literacy Rate") %>%
  hc_add_theme(hc_theme_google())

```


##### Mizoram has the highest Male Literacy and Female literacy rate in the country followed by Kerala.
##### Effective literacy rate of male increases as effective literacy rate of female increases. It has positive correlation



```{r}

hchart(by_state, "scatter", x = Male_Grads_Percent , y = Female_Grads_Percent, group = state_name) %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Male Graduate rate versus Female Graduate Rate") %>%
  hc_add_theme(hc_theme_google())

```


```{r}

hchart(by_state, "scatter", x = Male_Percent , y = Female_Percent, group = state_name) %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Male Population rate versus Female Population Rate") %>%
  hc_add_theme(hc_theme_google())

```


```{r}

hchart(indiancities, "scatter", x = sex_ratio, y = child_sex_ratio, group = state_name) %>%
  hc_credits(enabled = TRUE, text = "Sources: Census India 2011", style = list(fontSize = "10px")) %>%
  hc_title(text = "Sex Ratio versus Child Sex Ratio") %>%
  hc_add_theme(hc_theme_google())

```




##### Child sex ratio and Sex ratio has postive relation 




```{r}

sex_ratio_change <- indiancities %>% group_by(state_name) %>% summarise(Avg_Child_Ratio = mean(child_sex_ratio), Avg_Sex_Ratio = mean(sex_ratio)) %>% mutate(Diff = (Avg_Child_Ratio - Avg_Sex_Ratio ))

hchart(sex_ratio_change, x = state_name, y = Diff, type = "column", color = state_name) %>%
  hc_title(text = "Difference in Average Child Sex Ratio and Average Sex Ratio") %>%
  hc_add_theme(hc_theme_google())

```



```{r}

Diff_Child_Population <- indiancities %>% group_by(state_name) %>% summarise(Avg_Child_Population_Male = mean(X0.6_population_male), Avg_Child_Population_Female = mean(X0.6_population_female)) %>% mutate(Diff = (Avg_Child_Population_Male - Avg_Child_Population_Female))


hchart(Diff_Child_Population, x = state_name, y = Diff, type = "column", color = state_name) %>%
  hc_title(text = "Difference in Average Number of Male and Female Kids (Age Group 0-6)") %>%
  hc_subtitle(text = "Is this naturally possible or man made disaster?") %>%
  hc_add_theme(hc_theme_google())

```


```{r}

Diff_Avg_Effective_Literacy_Rate <- indiancities %>% group_by(state_name) %>% summarise(Avg_Effective_Literacy_Rate_Male = mean(effective_literacy_rate_male), Avg_Effective_Literacy_Rate_Female = mean(effective_literacy_rate_female)) %>% mutate(Diff = (Avg_Effective_Literacy_Rate_Male - Avg_Effective_Literacy_Rate_Female ))


hchart(Diff_Avg_Effective_Literacy_Rate, x = state_name, y = Diff, type = "column", color = state_name) %>%
  hc_title(text = "Difference in Average Effective Literacy Rate of Male and Female") %>%
  hc_subtitle(text = "Male Literacy Rate is higher in all the states") %>%
  hc_add_theme(hc_theme_google())

```


```{r}

Diff_Avg_graduates <- indiancities %>% group_by(state_name) %>% summarise(Avg_Graduates_Male = mean(male_graduates), Avg_Graduates_Female = mean(female_graduates)) %>% mutate(Diff = (Avg_Graduates_Male - Avg_Graduates_Female))


hchart(Diff_Avg_graduates, x = state_name, y = Diff, type = "column", color = state_name) %>%
  hc_title(text = "Difference in Average Number of Male and Female Graduates") %>%
  hc_subtitle(text = "Only Kerala has more Women Graduates than Male Graduates") %>%
  hc_add_theme(hc_theme_google())

```


```{r}

fivenum(indiancities$population_total)
quantile(indiancities$population_total)
quantile(indiancities$population_total)[5]/quantile(indiancities$population_total)[1]

hchart(indiancities$population_total) %>%
  hc_title(text = "Distirbution of the Total Population in Top 500 Cities") %>%
  hc_subtitle(text = "75% of the 500 Cities have less than 350K Population") %>%
  hc_add_theme(hc_theme_google())

```

#### 75% of the 500 Cities have less than 350K Population
#### Biggest city in Top 500 in terms of population is 124 times the smallest one.


```{r}

fivenum(indiancities$X0.6_population_total)
quantile(indiancities$X0.6_population_total)
quantile(indiancities$X0.6_population_total)[5]/quantile(indiancities$X0.6_population_total)[1]


hchart(indiancities$X0.6_population_total, color = "darkred") %>%
  hc_title(text = "Distirbution of the Total Population (0-6 Age Group) in Top 500 Cities") %>%
  hc_subtitle(text = "75% of the 500 Cities have less than 38K Population(0-6 Age Group") %>%
  hc_add_theme(hc_theme_google())

```


#### 75% of the 500 Cities have less than 38K Population
#### Biggest city in Top 500 in terms of population(0-6 age group) is 184 times the smallest one.


```{r}

fivenum(indiancities$literates_total)
quantile(indiancities$literates_total)
quantile(indiancities$literates_total)[5]/quantile(indiancities$literates_total)[1]

hchart(indiancities$literates_total, color = "darkgreen") %>%
  hc_title(text = "Distirbution of the Total Literates in Top 500 Cities") %>%
  hc_subtitle(text = "75% of the 500 Cities have less than 275K Literates") %>%
  hc_add_theme(hc_theme_google())

```


#### 75% of the 500 Cities have less than 275K Literates
#### Biggest city in Top 500 in terms of Number of Literates is 180 times the smallest one.
